---
title: "Stat 632 Group Project, Spring 2023"
author: "Linda, Siva, Brihana"
date: '2023-04-24'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(MASS)
library(car)
library(dplyr)
library(corrplot)
library(ggcorrplot)
library(ordinal)

#install.packages("corrplot")
```

```{r}
initial_data <- read.csv("ObesityDataSet_raw_and_data_sinthetic.csv", stringsAsFactors = T)

#physical_activity_data <- read.csv("Filtered_on_physcial_activity_Nutrition__Physical_Activity__and_Obesity_-_Behavioral_Risk_Factor_Surveillance_System_20230423.csv")
```


## Estimation of obesity levels based on eating habits and physical condition Data Set 

Paper: Dataset for estimation of obesity levels based on eating habits and physical condition in individuals from Colombia, Peru and Mexico
https://www.sciencedirect.com/science/article/pii/S2352340919306985#tbl1

Dataset
https://archive.ics.uci.edu/ml/datasets/Estimation+of+obesity+levels+based+on+eating+habits+and+physical+condition+


Displaying to the first 10 rows of the data set.
```{r}
head(initial_data, n = 10)
```


## Null Hypothesis:
Family history of obesity increases the risk of having obesity. 

## Alternative Hypothesis:
Family history does not increase the risk of having obesity. 



## Multiple Linear Regression Variables / Features:

@attribute Gender {Female,Male}
@attribute Age numeric
@attribute Height numeric
@attribute Weight numeric
@attribute family_history_with_overweight {yes,no}
@attribute FAVC {yes,no}
@attribute FCVC numeric
@attribute NCP numeric
@attribute CAEC {no,Sometimes,Frequently,Always}
@attribute SMOKE {yes,no}
@attribute CH2O numeric
@attribute SCC {yes,no}
@attribute FAF numeric
@attribute TUE numeric
@attribute CALC {no,Sometimes,Frequently,Always}
@attribute MTRANS {Automobile,Motorbike,Bike,Public_Transportation,Walking}
@attribute NObeyesdad {Insufficient_Weight,Normal_Weight,Overweight_Level_I,Overweight_Level_II,Obesity_Type_I,Obesity_Type_II,Obesity_Type_III}


## Data Prep


## Exploratory Data Analysis and Visualizations 

Summary of the data
```{r}
nrow(initial_data)
ncol(initial_data)
summary(initial_data)
```

## Initial correlation plot of the data
```{r}
num_cols <- sapply(initial_data, is.numeric)
mydata_num <- initial_data[, num_cols]
cor_mat <- cor(mydata_num)
corrplot(cor_mat)
```




## Checking for total number of NA's
```{r}
sum(is.na(initial_data))
```

## Plot of weight classification feature
```{r}
plot(initial_data$NObeyesdad)
```

## Converting categoricals factor to numeric
```{r}
## need to encode the factor variables 
new_data <- initial_data %>% mutate_if(is.factor, as.numeric)
head(new_data)
```
## Plotting the correlation plot.

```{r}
num_cols <- sapply(new_data, is.numeric)
mydata_num_1 <- new_data[, num_cols]
cor_mat <- cor(mydata_num_1)
corrplot(cor_mat)
```
## Scatter Plot Matrix of a Multiple Linear Regression Model

```{r}
pairs(new_data)
```
```{r}
view(new_data)
```



```{r}
#new_data_no_height_weight_df <- df[,c('Gender', 'Age', 'family_history_with_overweight', 'FAVC', 'FCVC', 'NCP', 'CAEC', 'SMOKE', 'CH2O', 'SCC', 'FAF', 'TUE', 'CALC', 'MTRANS')]

df_no_age_weight <- subset(new_data, select = -c(Height, Weight))
  #select(new_data, -Height, -Weight) 

head(df_no_age_weight)
```


## Converting the response variable into an ordinal. Testing this out after presentation was given. Commenting out.
```{r}
#df_no_age_weight$NObeyesdad <- ordered(df_no_age_weight$NObeyesdad, 
                                       #levels = c("Insufficient_Weight", 
                                      #            "Normal_Weight", 
                                      #            "Overweight_Level_I", 
                                      #            "Overweight_Level_II", 
                                      #            "Obesity_Type_I", 
                                      #            "Obesity_Type_II", 
                                      #            "Obesity_Type_III"))
```

```{r}
head(df_no_age_weight)
```

```{r}
summary(df_no_age_weight)
```

## Fitting the linear model.

```{r}
m1 <- lm(NObeyesdad ~ FCVC + NCP + CAEC + SMOKE + CH2O + SCC + FAF + TUE + CALC + MTRANS, data = df_no_age_weight)
```

## Summary of the model
`
```{r}
summary(m1)
```


```{r}
num_cols <- sapply(df_no_age_weight, is.numeric)
mydata_num_2 <- df_no_age_weight[, num_cols]
cor_mat_1 <- cor(mydata_num_2)
#M = corrplot(cor_mat)
corrplot(cor_mat_1, method = 'square', order = 'FPC', type = 'lower', diag = FALSE)
```

## Model fit, original (not converted to ordinal)

```{r}
m1 <- lm(NObeyesdad ~ Gender + Age + family_history_with_overweight + FAVC + FCVC + NCP  + CAEC + SMOKE + CH2O + SCC + FAF + TUE +CALC + MTRANS, data = new_data)
summary(m1)

```

```{r}
#m1 <- clm(NObeyesdad ~ FCVC + NCP + CAEC + SMOKE + CH2O + SCC + FAF + TUE + CALC + MTRANS, data = df_no_age_weight)
```

## Performing step function on the model using AIC: Akaike's An Information Criterion

```{r}
m2 <- step(m1)
```


## Creating a vector of model residuals
```{r}
residuals <- residuals(m2)
head(residuals)
```

```{r}
if (any(is.na(residuals) | !is.finite(residuals))) {
  stop("Residuals contain missing or infinite values. QQ plot cannot be created.")
}
```

# Create a QQ plot of the residuals
```{r}
qqnorm(residuals)
qqline(residuals)
```




## Summary of the model
```{r}
summary(m2)
```

## Residuals Plot of fitted model
```{r}
res <- resid(m2)
plot(fitted(m2), res)

#add a horizontal line at 0 
abline(0,0)
```

## Q-Q plot
```{r}
qqnorm(res)
qqline(res)
```

## Taking the log of the Weight value
```{r}
#new_data$Weight <- log(new_data$Weight)
#new_data$Weight
```


## Box-Cox to tranform data into a more normally distributed dataset.
```{r}
# library(MASS)
# library(car)
bcox <- boxcox(m2, lambda = seq(-5, 5, by = 0.1))
summary(powerTransform(m2))
```

```{r}
m3 <- lm(NObeyesdad^0.5247 ~., data=new_data)
m4 <- step(m3)
```

```{r}
summary(m4)
```


```{r}
AIC(m4)
```


```{}
regression_eq = -0.8377126 +  0.0040998(age) -0.5964036(height) + 2.1456942(weight) + 0.0323383(family_history)  -0.0618441(FAVC) -0.0464515(FCVC) -0.0275380(NCP) + 0.0387620(CAEC) + 0.0131555(CH2O) + 0.0787687(SCC) + 0.0062833(MTRANS)
```

## Partial f test, 0.3644
```{r}
anova(m3, m4)
```

## The value of p is greater than 0.05.So we fail to reject that the coefficients of these variables are equal to zero. This agrees with the step wise selection

```{r}
plot(m4)
```





